Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-03-30 arcticgrids.tex
new melt bias figure
paper completed; ready for coauthor comments

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}

@@

\correspondingauthor{=name=}{=email address=}

@@

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

@@

\begin{abstract}
[ enter your Abstract here ]
\end{abstract}

@@

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface. These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset. The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC. Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice. A unique surface energy balance and SMB is computed for each EC. Integrating over all ECs using the area weights provides a more accurate SMB. For a more detailed description of how the SMB is computed in CESM, the reader is referred to \cite{CISM1,SETAL2019TC,KETAL2020JAMES}.

@@

Several observational datasets are used in this study to understand the performance of the simulations. A list of the datasets used in this study are shown in Table~\ref{tbl:table2}. Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets were initially gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalyses at its domain boundaries. The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.

@@

\subsection{SMB Analysis}

A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area errors (computed as percent differences from the high resolution dataset) across the different grids, and as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.

The area errors in Figure~\ref{fig:grisdx} don't seem large, but even small changes in ice mask area can lead to large differences in integrated SMB \cite{HETAL2022TC}. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each dycore considered in this study. In addition, two remapping algorithms are used; Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative and TempestRemap \cite{TempestRemap} high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. ERA5 is very high-resolution at high-latitudes owing to the use of a $\frac{1}{4}^{\circ}$ latitude-longitude grid, and so all model differences are computed after mapping ERA5 to their respective model grids.}
\label{fig:dThyps}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset. The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.}
\label{fig:prect}
\end{figure}

@@

The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, expressed as the fractional difference from the RACMO2.3p2 solution. What sticks out is that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome. The $ARCTIC$ run improves this bias substantially, and the $ARCTICGIS$ run improves this bias further. This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.

Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast. These systems are orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast. At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping precipitation onto the interior of the ice sheet. A similar bias occurs in northwest Greenland, in particular during the summer, when it's common for synoptic systems to arrive from the southwest. The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point. Since the precipitation centers move from the interior towards the coasts, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly, eliminating this halo of low cloud bias around the oceanic perimeter of Greenland.

Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black. The 1979-1998 climatological mean values are provided as circles on the right side of the panels. The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates. The  variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO. The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr. The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).

The combined annual snow/ice melt integrated over the GrIS is shown in the bottom panel of Figure~\ref{fig:tseries}. The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO. The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr. This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}). The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, overestimating melt by only 70-90 Gt/yr. The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr. It is more difficult to attribute these differences to resolution alone, since the finite-volumes grids have colder summer temperatures than the uniform resolution spectral-element grids. However, that the $ARCTCIGRIS$ grid has the warmest summer temperatures, yet has a lower melting bias than the uniform spectral-element grids, suggests that increasing resolution improves the ablation process.

To get an idea of the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \cite{RM2012GRL}. The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented on the $f09$ and $ne30pg2$ grids (Figure~\ref{fig:melt}), which pertains to the grids in which all integrals are evaluated. Nonetheless, the southeast and southwest basins have the largest accumulation totals. All the basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions. The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins. Conversely, the $ARCTCIGIS$ precipitation is closest to the RACMO value in the large southwest basin.

The basin integrated melt rates in Figure~\ref{fig:basin} show a different grid sensitivity than the precipitation rates. Of note is that the uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins, with the $ARCTICGRIS$ grid a close second. This is somewhat unexpected as the $ARCTCIGRIS$ grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:dThyps}), yet it has less melting than the uniform-resolution spectral element grids. Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture local elevation effects as a result of a better representation of the low-elevation ablation zones. A more accurate proxy for the impact of temperature on melting is positive degree days \cite<PDD;>{B1984Z}. PDD is a metric that accumulates the near surface temperature in $C^{\circ}$ for any day in which the temperature is above freezing. PDD is non-linear in mean monthly near surface temperature \cite{R1991P}, and we compute it using monthly means 2-meter temperature using the method of \citeA{CG2005JG}.

The basin integrated PDD melt estimate is shown in Figure~\ref{fig:basin}. One interesting feature is that in the big southeast and southwest basin (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid. The southwest basin also predicts a rather large PDD based melt in the finite-volume grids, relative to the $ARCTCIGRIS$ grid. The PDD plots indicate that the near surface temperatures expected to contribute to melt is not well approximated by the summer lower troposphere temperatures in Figure~\ref{fig:dThyps}.

Figure~\ref{fig:melt} shows the biases in the combined ice/snow melt as map plots. This plot is quite revealing as it indicates that the largest melt biases occur on the southeast and northwest coast, where large coarse grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations then the ice sheet surface exists in nature, leading to too warm of a state and overestimating the melt. Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along two transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions (GMTED). The two transects are shown in Figure~\ref{fig:melt}; the infamous "k-transect" in southwest Greenland and a transect extending from the central dome down to the highly productive Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $f09$ grid is a bit of an exception; the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect. The variable-resolution grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of ice sheets.

While the two transects in Figure~\ref{fig:ztrans} only two anecdotes, they indicate that the ice sheet surface is not systematically lower than the true surface in the coarse grids, near the ice margins. Rather the smoothing of the raw topography, necessary to prevent coarse models from instigating grid-scale models, makes the ice sheet margin more gradual, causing the lower-elevation ablation zones to extend further away and outside the true ice sheet margin, resulting in the coarse cells in the ablation zone situated higher than the actual ice surface. This is another clue; the overextending of the ice sheet out over the ice margin means that these grid cells have multiple surface types within them. That is, It is these mixed ice/land/ocean grid cells tend to have the largest positive melt bias. It seems likely based on the transect that some of these largely biased coarse cells only have small fractions of ice, and so the state of the grid cell may be predominantly controlled by the land/ocean surface types.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccc}
   \hline
   grid name & accumulation (Gt/yr) & ice/snow melt (Gt/yr) \\ 
   \hline
   $RACMO$ & 768.5 & -347.2 \\
   \hline
   $f19$ & 882.5 & -440.3 \\
   $f09$ & 874.8 & -418.4 \\
   $ne30pg2$ & 1000. & -549.4 \\
   $ne30pg3$ & 934.9 & -568.8 \\
   $ARCTIC$ & 795.9 & -367.3 \\
   $ARCTICGRIS$ & 708.7 & -471.6 \\
 \hline
 \end{tabular}
  \caption{1979-1998 Surface Mass Balance of the Greenland Ice Sheet.}
 \label{tbl:table3}
 \end{table*}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
\end{center}
\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to RACMO3.2. The raw fields are mapped to two target low resolution grids, f19 \& ne30pg2, and using two different remapping methods, conservative ESMF and high order TempestRemap. The remapped values are then integrated over the ice mask of the target grid. This gives four time-series for each simulation (three for f19 \& ne30pg2), with the mean value given by the solid line and shading spanning the extent of the remapped solutions.}
\label{fig:tseries}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
\end{center}
\caption{.}
\label{fig:basin}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
\end{center}
\caption{.}
\label{fig:melt}
\end{figure}

Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots. The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates. The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller. The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, providing confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone. The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}. They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, and that the $ARCTICGRIS$ performs best, and an improvement over the $ARCTIC$ grid. However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.

%\begin{figure}[t]
%\begin{center}
%         \includegraphics[width=130mm]{figs/temp_violens.pdf}
%\end{center}
%\caption{.}
%\label{fig:pointdiff}
%\end{figure}

>>>>

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, Marcus Lofverstrom\affil{2}, Andrew Gettelman\affil{1}}

@@

\correspondingauthor{Adam Herrington}{aherring@ucar.edu}

@@

\begin{keypoints}
\item The transition from latitude-longitude to unstructured grids in CESM negatively impacts the surface mass balance of the Greenland Ice Sheet.
\item Two new Arctic refined mesh configurations have been developed and made available in CESM2.2
\item The Arctic meshes provide improvements to the surface mass balance over all $1^{\circ}$ grids, including latitude-longitude grids.
\end{keypoints}

@@

\begin{abstract}

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both the accumulation and ablation processes over the GrIS. Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS. Two Arctic refined-meshes, containing $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented in this study as newly supported configurations in CESM2.2. The Arctic meshes provide substantial improvements to the simulated cloud field in the Arctic and over Greenland, resulting in a skillful representation of the accumulation and ablation processes, and a more realistic GrIS surface mass balance compared with the conventional, uniform resolution grids. As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes serve to recover, and improve upon the representation of polar processes in CESM.

\end{abstract}

@@

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation runs-off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, representing an accumulation of mass. These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset. The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC. Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice. A unique surface energy balance and SMB is computed for each EC. Integrating over all ECs using the area weights provides a more accurate SMB. For a more detailed description of how the SMB is computed in CESM, the reader is referred to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES}.

@@

Several observational datasets are used in this study to gauge the performance of the simulations. A list of the datasets used in this study are provided in Table~\ref{tbl:table2}. Several of these products are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets were initially gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim and ERA5 renalysis products at its domain boundaries. The RACMO simulations have been shown to perform skillfully against observations and is often used as an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.

@@

\subsection{SMB Analysis}\label{sec:SMB}

We seek to integrate components of the SMB over a GrIS ice mask and diagnose their contributions to the mass budget of the GrIS. However, the ice masks vary across all the grids, especially in comparison to the RACMO3.2 ice mask which is about $3\%$ smaller than the reference dataset (Figure~\ref{fig:grisdx}). The area errors in Figure~\ref{fig:grisdx} may not seem large, but even order $1-2\%$ percent ice mask area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}. CLM's dataset creation tools generates the model ice mask by mapping a raw high resolution dataset containing the GrIS boundary conditions, to the target grid using the Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative remapping algorithm. The figure suggests the mapping errors are less than $1.5\%$ across just the model grids.

In this study we have taken a common ice mask approach by mapping all model fields to the lowest resolution grids and integrating over their respective low resolution ice masks. A low resolution common ice mask is a conservative decision and is largely because we seek to use first-order remapping algorithms for mapping fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher resolution grid than the source grid. Two remapping algorithms are used; ESMF first-order conservative and TempestRemap \cite{TempestRemap} high-order monotone algorithm. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each grid coordinate type. In all, each integrated quantity is evaluated (at most) four times per grid to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

We consider two different approaches for integrating mapped fields over the common ice mask.
\begin{enumerate}
\item Map the patch level quantities, or the state over the ice fractional component of the grid cell, to the common grid, and integrate the mapped field over the common ice mask. \label{lbl:label1}
\item Map the grid cell mean quantities to the common grid, and integrate the mapped field over the common ice sheet. \label{lbl:label2}
\end{enumerate}

Method~\ref{lbl:label1} introduces errors that is related to differences in the ice mask between the source and target grid. Since the lower resolution ice masks tend to have larger GrIS areas (Figure~\ref{fig:grisdx}), we expect these errors to inflate the resulting integrals. Method~\ref{lbl:label2} assumes the patch level values are identical to the grid mean values. This has the effect of homogenizing ice mask specific states over the entire grid cell. This is especially detrimental to the ice melt quantity, since ice melt can only occur over the ice mask. The accumulation term (combined solid plus liquid precipitation), however, should not be adversely impacted by method~\ref{lbl:label2} since it is not directly tied to the ice mask, but instead controlled by larger resolved scales. One could also argue that method~\ref{lbl:label1} is less accurate for the accumulation term, since it does not map to the new ice mask as cleanly as in method~\ref{lbl:label2}.

Method~\ref{lbl:label2} may be necessary in situations where only the grid mean values are available, and the patch level values cannot be reconstructed. This is unfortunately the case for the RACMO datasets; ice melt cannot be reconstructed at the ice patch level. Method~\ref{lbl:label2} is therefore used in this study, and we should expect a dissipative effect when diagnosing the melt rates. As we will show later, this method still preserves the relative differences between the different grids and in comparison with the RACMO datasets.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $f09$ grid, inline.}
\label{fig:dThyps}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset. The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period. \color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
\label{fig:prect}
\end{figure}

@@

The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:prect}, expressed as the fractional difference from the RACMO2.3p2 solution. It is evident that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome. The $ARCTIC$ run improves this bias substantially, and the $ARCTICGRIS$ run improves this bias further. This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.

Large accumulation rates in GrIS result from synoptic systems arriving from the south. These systems are orographiclly lifted at the ice sheet margin, especially the steep slopes in southeast Greenland, concentrating large amounts of precipitation about the ice margin. At lower horizontal resolutions the topography is too smooth and anomalous moisture penetrates inland, erroneously dumping precipitation onto the southern dome. The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point. As the precipitation centers move from the interior towards the margins, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly. Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, \ref{fig:prect} clearly illustrate a reduction in cloud amount in the interior and an increased cloudiness around the oceanic perimeter of Greenland, in comparing the uniform $1^{\circ}-2^{\circ}$ runs to the variable-resolution runs.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
   \hline
   $RACMO$ & 768.5 (733.5) & -347.2 (-436.4) & -221.7 (-258.5) & -36.5 (-38.8) & 510.3 (436.2) \\
   \hline
   $f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
   $f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
   $ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
   $ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
   $ARCTIC$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
   $ARCTICGRIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
 \hline
 \end{tabular}
  \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr. Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
 \label{tbl:table3}
 \end{table*}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
\end{center}
\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets. The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
\label{fig:tseries}
\end{figure}

Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components in all the grids, and compared with RACMO. Note the values in the table averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}. The RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common ice mask approach. The SMB is computed from the individual components as:
\begin{equation}
SMB = accumulation + runoff + sublimation \label{eq:SMB}
\end{equation}
and all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass). Note that this glaciological SMB differs from the internally computed SMB in CLM; see \citeA{KETAL2020JAMES} for an explanation of the differences.

Table~\ref{tbl:table3} also contain's (in parentheses) the SMB components derived from evaluating the integrals on each model's native grid and ice mask. Of note is the large reduction in ice/melt rates using the common ice mask approach compared with that computed on the native grid. This illustrates the dissipative impact of the common ice mask approach discussed in section~\ref{sec:SMB}. The errors are large because it is mixed ice/ice-free grid cells straddling the ice sheet margins that are more prone to errors, which are commonly in the ablation zone and have large melt rates. Contrast this with the differences in integrated precipitation, where the differences between the native approach and common ice mask approach is much less. As discussed in the methods, this is because the combined solid/liquid precipitation rates are not directly tied to the ice mask, whereas ice melt is exclusively tied to the ice mask.

Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black. The 1979-1998 climatological mean values, tabulated in Table~\ref{tbl:table3}, are also provided as circles on the right side of the panels. The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates. The variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO. The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr. The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).

The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO. The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr. This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}). The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, overestimating melt by only 70-90 Gt/yr. The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr. It is more difficult to attribute these differences to resolution alone, since the finite-volume grids have colder summer temperatures than the uniform resolution spectral-element grids.

To understand the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}. The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which can be seen in the $f19$ and $ne30pg2$ panels of Figure~\ref{fig:melt}. Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the largest accumulation totals. All basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions. The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation amounts in the large southwest basin.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
\end{center}
\caption{1979-1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method. Whiskers span the max/min of the 4 ensemble members generated from the common ice mask approach. Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:melt}.}
\label{fig:basin}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
\end{center}
\caption{1979-1998 ice/snow melt biases (in mm/yr) relative to RACMO2.3p2, evaluated on the native model grids. The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $f19$ and $ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
\label{fig:melt}
\end{figure}

The basin integrated melt rates in Figure~\ref{fig:basin} show a dycore dependent-type pattern. The uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins. The $ARCTICGRIS$ grid is a close second, while the finite-volume grids have systematically smaller melt-rates. The ``second-place" standing of $ARCTICGRIS$ is somewhat unexpected as this grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution spectral element grids.

Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones. A more accurate proxy for the impact of temperature on melting is positive degree-days \cite<PDD;>{B1984Z}. PDD is a metric that accumulates the near surface temperature in $^{\circ}C$ for any day in which the temperature is above freezing. PDD is non-linear in mean monthly temperature \cite{R1991P}, and we compute it using monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}C$, and a degree-day factor of 5 $mm \ d^{-1}$ $^{\circ}C^{-1}$.

The basin integrated PDD melt estimate is shown in Figure~\ref{fig:basin}. One interesting feature is that in the big southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid. The southwest basin also predicts a rather large PDD based melt in the finite-volume grids, relative to the $ARCTCIGRIS$ grid. The PDD plots indicate that the near surface temperatures expected to contribute to melt is not well approximated by the summer lower troposphere temperatures in Figure~\ref{fig:dThyps}.
